---
title: "Conrad V3 SDA Sensitivity Analysis"
output:
  html_notebook: 
    code_folding: hide
    fig_height: 6
    fig_width: 10.5
    results: hold
    fig_show: hold
    toc: yes
    toc_float: yes
    number_sections: yes
    theme: spacelab
---

SDA is deterministic given a random starting configuration, so we test multiple random seeds to ensure stability of the results.

# Baseline

This is the result used in the rest of the analysis and we will compare the results of other runs to this.

We will compare two components per run, a large one (pachytene - marked by Nasp), and a smaller one (leptotene - marked by Prdm9).


```{r}
library(SDAtools)

run_SDA(sda_location = "../../SDA/build/sda",
        out = "../results/conradV3_sda_1",
        data = "V3_SDAmerged_mouse_V3_SDA.data",
        num_comps = 50,
        max_iter = 10000,
        save_freq = 1000,
        set_seed = "79151 17351",
        N = 20322)

results1 <- load_results(results_folder = "../data/conradV3/conradV3_sda_1/", data_path = "../data/conradV3/")
rownames(results1$loadings[[1]]) <- paste0("V",1:50)
rownames(results1$pips[[1]]) <- paste0("V",1:50)

str(results1)
max(results1$free_energy)
check_convergence(results1)

highest_components(results1, "Nasp")
highest_genes(results1, component = 42)

highest_components(results1, "Prdm9")
highest_genes(results1, component = 5)
```

# Different Seed 1

```{r}
# try different seeds to test consistancy
run_SDA(sda_location = "../../SDA/build/sda",
        out = "../results/conradV3_sda_5",
        data = "../data/conradV3/V3_SDAmerged_mouse_V3_SDA.data",
        num_comps = 50,
        max_iter = 10000,
        save_freq = 1000,
        set_seed = "82145 34369",
        N = 20322)

results5 <- load_results(results_folder = "../data/conradV3/conradV3_sda_5/", data_path = "../data/conradV3/")
rownames(results5$loadings[[1]]) <- paste0("V",1:50)
rownames(results5$pips[[1]]) <- paste0("V",1:50)

str(results5)
max(results5$free_energy)
check_convergence(results5)
loading_distribution(results5)

highest_components(results5, "Nasp")
plot(results1$loadings[[1]][42,], results5$loadings[[1]][5,])
highest_genes(results5, component = 5)

highest_components(results5, "Prdm9")
plot(results1$loadings[[1]][5,], results5$loadings[[1]][26,])
highest_genes(results5, component = 26)
```

# Different Seed 2

```{r}
run_SDA(sda_location = "../../SDA/build/sda",
        out = "../results/conradV3_sda_6",
        data = "../data/conradV3/V3_SDAmerged_mouse_V3_SDA.data",
        num_comps = 50,
        max_iter = 10000,
        save_freq = 1000,
        set_seed = "69808 6210",
        N = 20322)

results6 <- load_results(results_folder = "../data/conradV3/conradV3_sda_6/", data_path = "../data/conradV3/")
rownames(results6$loadings[[1]]) <- paste0("V",1:50)
rownames(results6$pips[[1]]) <- paste0("V",1:50)

str(results6)
max(results6$free_energy)
check_convergence(results6)

highest_components(results6, "Nasp")
plot(results1$loadings[[1]][42,], results6$loadings[[1]][45,])
highest_genes(results6, component = 45)

highest_components(results6, "Prdm9")
plot(results1$loadings[[1]][5,], results6$loadings[[1]][49,])
highest_genes(results6, component = 49)
```

# Different Seed 3

```{r}
run_SDA(sda_location = "../../SDA/build/sda",
        out = "../results/conradV3_sda_7",
        data = "../data/conradV3/V3_SDAmerged_mouse_V3_SDA.data",
        num_comps = 50,
        max_iter = 10000,
        save_freq = 1000,
        set_seed = "31111 30486",
        N = 20322)

results7 <- load_results(results_folder = "../data/conradV3/conradV3_sda_7/", data_path = "../data/conradV3/")
rownames(results7$loadings[[1]]) <- paste0("V",1:50)
rownames(results7$pips[[1]]) <- paste0("V",1:50)

str(results7)
max(results7$free_energy)
check_convergence(results7)

highest_components(results7, "Nasp")
plot(results1$loadings[[1]][42,], results7$loadings[[1]][10,])
highest_genes(results7, component = 10)

highest_components(results7, "Prdm9")
plot(results1$loadings[[1]][5,], results7$loadings[[1]][5,])
highest_genes(results7, component = 5)
```

# Different Seed 4

```{r}
run_SDA(sda_location = "../../SDA/build/sda",
        out = "../results/conradV3_sda_8",
        data = "../data/conradV3/V3_SDAmerged_mouse_V3_SDA.data",
        num_comps = 50,
        max_iter = 10000,
        save_freq = 1000,
        set_seed = "79874 65015",
        N = 20322)

results8 <- load_results(results_folder = "../data/conradV3/conradV3_sda_8/", data_path = "../data/conradV3/")
rownames(results8$loadings[[1]]) <- paste0("V",1:50)
rownames(results8$pips[[1]]) <- paste0("V",1:50)

str(results8)
max(results8$free_energy)
check_convergence(results8)

highest_components(results8, "Lyar")
plot(results1$loadings[[1]][42,], results8$loadings[[1]][39,])
highest_genes(results8, component = 39)

highest_components(results8, "Prdm9")
plot(results1$loadings[[1]][5,], results8$loadings[[1]][30,])
highest_genes(results8, component = 30)

```

# 200 Components

```{r}
# try with different seed & higher components
# higher components
# run_SDA(sda_location = "../../SDA/build/sda",
#         out = "../results/conradV3_sda_2",
#         data = "../data/conradV3/V3_SDAmerged_mouse_V3_SDA.data",
#         num_comps = 75,
#         max_iter = 10000,
#         save_freq = 1000,
#         set_seed = "73453 98253",
#         N = 20322,
#         eigen_parallel = TRUE,
# 	      num_blocks = 128,
# 	      num_openmp_threads = 12)

# Even more components
run_SDA(sda_location = "../../SDA/build/sda",
        out = "../results/conradV3_sda_3",
        data = "../data/conradV3/V3_SDAmerged_mouse_V3_SDA.data",
        num_comps = 200,
        max_iter = 5000,
        save_freq = 500,
        set_seed = "79151 17351",
        N = 20322,
        eigen_parallel = TRUE,
	      num_blocks = 128,
	      num_openmp_threads = 12)

results3 <- load_results(results_folder = "../data/conradV3/conradV3_sda_3/", data_path = "../data/conradV3/")
rownames(results3$loadings[[1]]) <- paste0("V",1:200)
rownames(results3$pips[[1]]) <- paste0("V",1:200)

str(results3)
max(results3$free_energy)
check_convergence(results3)

highest_components(results3, "Lyar")
plot(results1$loadings[[1]][42,], results3$loadings[[1]][157,])
highest_genes(results3, component = 157)

highest_components(results3, "Prdm9")
plot(results1$loadings[[1]][5,], results3$loadings[[1]][138,])
highest_genes(results3, component = 138)

```

# Centered asinh

Using a different normalisation Asinh is like log when x is large but works when x=0

```{r}
library(data.table)
library(ggplot2)
tmp <- data.table(x=seq(-15,50,0.01),
                  log=log(seq(-15,50,0.01)),
                  log1p=log1p(seq(-15,50,0.01)),
                  asinh=asinh(seq(-15,50,0.01)),
                  sqrt=sqrt(seq(-15,50,0.01)))

ggplot(melt(tmp[x<20], id.vars = "x"), aes(x, value, colour=variable)) +geom_line() + theme_minimal() + geom_abline(intercept = 0,slope=1) + ylim(0,NA) + xlim(0,NA)
```



```{r}

export_data(as.matrix(data), name = "merged_mouse_V3_SDA_asinh_c", path="../data/conradV3/")
run_SDA(sda_location = "../../SDA/build/sda",
        out = "../results/conradV3_sda_4",
        data = "../data/conradV3/merged_mouse_V3_SDA_asinh_c.data",
        num_comps = 50,
        max_iter = 5000,
        save_freq = 500,
        set_seed = "79151 17351",
        N = 20322,
        eigen_parallel = TRUE,
	      num_blocks = 128,
	      num_openmp_threads = 12)

results4 <- load_results(results_folder = "../data/conradV3/conradV3_sda_4/", data_path = "../data/conradV3/")
rownames(results4$loadings[[1]]) <- paste0("V",1:50)
rownames(results4$pips[[1]]) <- paste0("V",1:50)

str(results4)
max(results4$free_energy)
check_convergence(results4)

highest_components(results4, "Lyar")
plot(results1$loadings[[1]][42,], results4$loadings[[1]][47,])
highest_genes(results4, component = 47)

highest_components(results4, "Prdm9")
plot(results1$loadings[[1]][5,], results4$loadings[[1]][49,])
highest_genes(results4, component = 49)

```


