---
title: "Individual Components"
output: 
  html_notebook: 
    code_folding: hide
    fig_height: 6
    fig_width: 10.5
    results: hold
    fig_show: hold
    toc: yes
    toc_float: yes
    number_sections: yes
    theme: spacelab
---

```{r}
library(DT)
motifEnrichment_wLogo <- readRDS("../data/motifs/TFenrichment_V2.rds")
GO_data <- readRDS("../data/GO_enrichment.rds")

library(testisAtlas)
load("../data/tmp_cache/SDA_objects.rds")
load_component_orderings()
```

# Odd / Unassigned Components

## 22 Odd

```{r, fig.width=6}
print_tsne(22, point_size = 0.5)
print_loadings_scores(22)
print_gene_list(22)

go_volcano_plot(component = "V22N")
head(GO_data[["V22N"]][c(2,7,8,10)], 20)

datatable(motifEnrichment_wLogo[geneSet=="V22N",-c("enrichedGenes","geneSet"), with=FALSE][1:40,], escape = FALSE, filter="top", options=list(pageLength=5))
          
```

## 6 Mj

```{r, fig.width=6}
print_tsne(6, point_size = 0.5)
print_loadings_scores(6)
print_gene_list(6)

go_volcano_plot(component = "V6P")
head(GO_data[["V6P"]][c(2,7,8,10)], 20)

datatable(motifEnrichment_wLogo[geneSet=="V6P",-c("enrichedGenes","geneSet"), with=FALSE][1:40,], escape = FALSE, filter="top", options=list(pageLength=5))
          
```


## 24 Leydig Subset

Loading appears in a subset of leydig cells, but alot of the top (positive) genes are pseudogenes, and most almost all high cell scores are in Hormad1 KO so maybe a batch / KO specific effect / messed up.

```{r, fig.width=6}
print_tsne(24, point_size = 0.5)
print_loadings_scores(24)
print_gene_list(24)

go_volcano_plot(component = "V24N")
head(GO_data[["V24N"]][c(2,7,8,10)], 20)

head(GO_data[["V24P"]][c(2,7,8,10)], 20)

datatable(motifEnrichment_wLogo[geneSet=="V24N",-c("enrichedGenes","geneSet"), with=FALSE][1:40,], escape = FALSE, filter="top", options=list(pageLength=5))
          
```




## 29 SPG batch??

Spermatogonial location, but ribosomal negative genes, and top genes all non-protein coding or mitochondrial:

- Gm14513 - pseudogene
- Rps15a-ps8 - pseudogene
- Gm15772 - pseudogene
- Gm14094 - pseudogene
- Gm44067 - lincRNA
- Gm13092 - pseudogene
- Gm10086 - pseudogene
- AC129329.1 - pseudogene

```{r, fig.width=6}
print_tsne(29, point_size = 0.5)
print_loadings_scores(29)
print_gene_list(29)

go_volcano_plot(component = "V29P")
head(GO_data[["V29P"]][c(2,7,8,10)], 20)
head(GO_data[["V29N"]][c(2,7,8,10)], 20)

datatable(motifEnrichment_wLogo[geneSet=="V25N",-c("enrichedGenes","geneSet"), with=FALSE][1:40,], escape = FALSE, filter="top", options=list(pageLength=5))
          
```






## 12 Batch (Hormad1)

Odd tSNE pattern, cells on outside + blob in center. One set of Hormad1 KO cells have opposite cell loadings to the others so seems unlikely due to the KO of Hormad1 and more due to a batch? - but also high scores in SPG FACS. Highest gene (Gm42418) is lincRNA, mt-Rnr1 & mt-Rnr2 also high.

```{r, fig.width=6}
print_tsne(12, point_size = 0.5)
print_loadings_scores(12)
print_gene_list(12)

go_volcano_plot(component = "V12N")
head(GO_data[["V12N"]][c(2,7,8,10)], 20)

datatable(motifEnrichment_wLogo[geneSet=="V12N",-c("enrichedGenes","geneSet"), with=FALSE][1:40,], escape = FALSE, filter="top", options=list(pageLength=5))
          
```

## 28 Acrosome V2 / CUl4a Batch

Overlaps V30, potentially batch effect as top 15 positive genes only 2 are protein coding, the rest are either lincRNA, antisenseRNA or processes_pseudogene.

```{r, fig.width=6}
print_tsne(28, point_size = 0.5)
print_loadings_scores(28)
print_gene_list(28)

head(GO_data[["V28P"]][c(2,7,8,10)], 20)
head(GO_data[["V28N"]][c(2,7,8,10)], 20)

datatable(motifEnrichment_wLogo[geneSet=="V28N",-c("enrichedGenes","geneSet"), with=FALSE][1:40,], escape = FALSE, filter="top", options=list(pageLength=5))
          
```

## 36 Spermatogenesis

Small number of cells on the outside, cells at spermatogenesis stage in tsne, but negative gene loadings (matching negative cell scores) have no enrichment & a few lincRNAs

```{r, fig.width=6}
print_tsne(36, point_size = 0.5)
print_loadings_scores(36)
print_gene_list(36)

go_volcano_plot(component = "V36P")
head(GO_data[["V36P"]][c(2,7,8,10)], 20)
head(GO_data[["V36N"]][c(2,7,8,10)], 20)

datatable(motifEnrichment_wLogo[geneSet=="V36P",-c("enrichedGenes","geneSet"), with=FALSE][1:40,], escape = FALSE, filter="top", options=list(pageLength=5))
          
```


## 41 batch

looks clumpy in tsne, potentially WT batch effect

```{r, fig.width=6}
print_tsne(41, point_size = 0.5)
print_loadings_scores(41)
print_gene_list(41)

go_volcano_plot(component = "V41P")
head(GO_data[["V41P"]][c(2,7,8,10)], 20)

datatable(motifEnrichment_wLogo[geneSet=="V41P",-c("enrichedGenes","geneSet"), with=FALSE][1:40,], escape = FALSE, filter="top", options=list(pageLength=5))
          
```

## 9 Respiration

```{r, fig.width=6}
print_tsne(9, point_size = 0.5)
print_loadings_scores(9)
print_gene_list(9)

go_volcano_plot(component = "V9N")
head(GO_data[["V9N"]][c(2,7,8,10)], 20)

head(GO_data[["V9P"]][c(2,7,8,10)], 20)

datatable(motifEnrichment_wLogo[geneSet=="V9N",-c("enrichedGenes","geneSet"), with=FALSE][1:40,], escape = FALSE, filter="top", options=list(pageLength=5))
          
```




## 49 ?? Sertoli / Leydig?

```{r, fig.width=6}
print_tsne(49, point_size = 0.5)
print_loadings_scores(49)
print_gene_list(49)

go_volcano_plot(component = "V49P")
head(GO_data[["V49P"]][c(2,7,8,10)], 20)

go_volcano_plot(component = "V49N")
head(GO_data[["V49N"]][c(2,7,8,10)], 20)


datatable(motifEnrichment_wLogo[geneSet=="V49P",-c("enrichedGenes","geneSet"), with=FALSE][1:40,], escape = FALSE, filter="top", options=list(pageLength=5))
          
```

## 43 ribosomal

```{r, fig.width=6}
print_tsne(43, point_size = 0.5)
print_loadings_scores(43)
print_gene_list(43)

go_volcano_plot(component = "V43N")
head(GO_data[["V43N"]][c(2,7,8,10)], 20)

datatable(motifEnrichment_wLogo[geneSet=="V43N",-c("enrichedGenes","geneSet"), with=FALSE][1:40,], escape = FALSE, filter="top", options=list(pageLength=5))
          
```



Cxadr aka CAR [Coxsackie and adenovirus receptor (CAR) is a product of Sertoli and germ cells in rat testes which is localized at the Sertoli-Sertoli and Sertoli-germ cell interface](https://dx.doi.org/10.1016%2Fj.yexcr.2007.01.017)

## 48 Odd clump

Highest in Mlh3? Odd location on tsne.

```{r, fig.width=6}
print_tsne(48, point_size = 0.5)
print_loadings_scores(48)
print_gene_list(48)

go_volcano_plot(component = "V48P")
head(GO_data[["V48P"]][c(2,7,8,10)], 20)

datatable(motifEnrichment_wLogo[geneSet=="V48P",-c("enrichedGenes","geneSet"), with=FALSE][1:40,], escape = FALSE, filter="top", options=list(pageLength=5))
          
```

## 1 single

A single blood cell? / bad component

```{r, fig.width=6}
print_tsne(1, point_size = 1)
print_loadings_scores(1)
#print_gene_list(1)

#go_volcano_plot(component = "V1P")
head(GO_data[["V1P"]][c(2,7,8,10)], 20)

#datatable(motifEnrichment_wLogo[geneSet=="V1P",-c("enrichedGenes","geneSet"), with=FALSE][1:40,], escape = FALSE, filter="top", options=list(pageLength=5))
```

## 46 single

```{r, fig.width=6}
print_tsne(46, point_size = 1)
print_loadings_scores(46)
#print_gene_list(46)

#go_volcano_plot(component = "V46N")
head(GO_data[["V46N"]][c(2,7,8,10)], 20)

#datatable(motifEnrichment_wLogo[geneSet=="V46N",-c("enrichedGenes","geneSet"), with=FALSE][1:40,], escape = FALSE, filter="top", options=list(pageLength=5))
          
```

## 4 Single
```{r, fig.width=6}
print_tsne(4, point_size = 1)
print_loadings_scores(4)
#print_gene_list(4)

#go_volcano_plot(component = "V4N")
head(GO_data[["V4N"]][c(2,7,8,10)], 20)

#datatable(motifEnrichment_wLogo[geneSet=="V4N",-c("enrichedGenes","geneSet"), with=FALSE][1:40,], escape = FALSE, filter="top", options=list(pageLength=5))
          
```

## 8 Single
```{r, fig.width=6}
print_tsne(8, point_size = 1)
print_loadings_scores(8)
#print_gene_list(8)

#go_volcano_plot(component = "V8N")
head(GO_data[["V8N"]][c(2,7,8,10)], 20)

#datatable(motifEnrichment_wLogo[geneSet=="V8N",-c("enrichedGenes","geneSet"), with=FALSE][1:40,], escape = FALSE, filter="top", options=list(pageLength=5))
          
```

## 14 Single

```{r, fig.width=6}
print_tsne(14, point_size = 1)
print_loadings_scores(14)
#print_gene_list(14)

#go_volcano_plot(component = "V14P")
head(GO_data[["V14P"]][c(2,7,8,10)], 20)

#datatable(motifEnrichment_wLogo[geneSet=="V14P",-c("enrichedGenes","geneSet"), with=FALSE][1:40,], escape = FALSE, filter="top", options=list(pageLength=5))
          
```
