---
title: "Systematic Denovo Motif Finding"
output: 
  html_notebook: 
    code_folding: hide
    results: hold
    fig_show: hold
    toc: yes
    toc_float: yes
    number_sections: yes
    theme: spacelab
---

# De Novo Motif Finding

Find denovo motifs for every component, using varying parameters -> 9000 potential motifs total.

Setup
```{r}
library(MotifFinder)
library(SDAtools)
library(ggplot2)
library(ggseqlogo)
nmf.options(grid.patch=TRUE) # stop blank pages appearing

library(testisAtlas)
#_500
tss_seqs <- seqinr::read.fasta("../data/motifs/promoter_sequences/all_genes_tss_1000.fasta", forceDNAtolower = FALSE, as.string = TRUE)
#motifs/promoter_sequences
tss_seqs_s <- gsub("a|t|c|g|n","N",as.character(tss_seqs))
names(tss_seqs_s) <- gsub("\\(-\\)|\\(\\+\\)","",names(tss_seqs))

str(tss_seqs_s)

SDAresults <- load_results(results_folder = "../data/conradV3/conradV3_sda_1/", data_path = "../data/conradV3/")
rownames(SDAresults$loadings[[1]]) <- paste0("V",1:50)

```


Run in parallel

```{r}
# test on a single component
#test <- get_component_motifs(5,T)

results_negative <- mclapply(1:50, function(x) get_component_motifs(x, FALSE), mc.cores = 7, mc.silent=TRUE, mc.preschedule = FALSE)

results_positive <- mclapply(1:50, function(x) get_component_motifs(x, TRUE), mc.cores = 7, mc.silent=TRUE, mc.preschedule = FALSE)


# results_random <- mclapply(1:50, function(x) get_component_motifs(x, positive=FALSE, random=TRUE), mc.cores = 15, mc.silent=TRUE, mc.preschedule = FALSE)

# a couple errored due to NA in creating seqs regs e.g. 14? / a single cell one

saveRDS(results_negative, "motifFinder_denovo_results_negative_V3.rds")

saveRDS(results_positive, "motifFinder_denovo_results_positive_V3.rds")

# saveRDS(results_random, "motifFinder_denovo_results_random.rds")

```


Clean up & Combine results.

```{r}


results_positive <- readRDS("motifFinder_denovo_results_positive_V3.rds")

table(unlist(lapply(results_negative, length)))
# expect 50 components with length 90

table(unlist(lapply(results_positive, length)))

table(unlist(lapply(results_random, length)))

# Add names to each motif found

names(results_positive) <- paste0("C",1:50,"P")
names(results_negative) <- paste0("C",1:50,"N")
names(results_random) <- paste0("C",1:50,"R")

for(i in 1:50){
  names(results_positive[[i]]) <- paste0("I",1:90)
  names(results_negative[[i]]) <- paste0("I",1:90)
#  names(results_random[[i]]) <- paste0("I",1:90)
}

results_all <- c(unlist(results_positive, recursive = FALSE),
                 unlist(results_negative, recursive = FALSE))

sum(sapply(results_all, is.null))
#V3: 1427 failed / 9000

# above is rather large due to sequences kept, so remove
for(i in 1:length(results_all)){
  if(!is.null(results_all[[i]])){
    results_all[[i]]$seqs <- names(results_all[[i]]$seqs)
    results_all[[i]]$trimmedseqs <- NULL  
  }
}


saveRDS(results_all, "motifFinder_denovo_results_all_clean_V3.rds")

# results_all_random <- unlist(results_random, recursive = FALSE)
# 
# saveRDS(results_all_random, "motifFinder_denovo_results_all_random_clean.rds")
```

# Find motif matches in HOCOMOCO using TOMTOM

Export denovo motifs in MEME format to input to TomTom to match to known motifs

```{r}

system("mkdir all_denovo_pwms_V3")

export_all_pwms(results_all, "all_denovo_pwms_V3/all_denovo_pwms_V3.meme")

# export_all_pwms(results_all_random, "random_pwms.meme")

```

Run tomtom

```{bash, eval=FALSE}
wget http://meme-suite.org/meme-software/5.0.4/meme-5.0.4.tar.gz

tar zxf meme-5.0.4.tar.gz
cd meme-5.0.4
./configure --prefix=$HOME/meme --with-url=http://meme-suite.org --enable-build-libxml2 --enable-build-libxslt
make
make test
make install

mkdir all_denovo_PWM_tomtom_v3

# Run TOMTOM in parallel (it's a bit slow)

for batch in 0 1 2 3 4 5 6 7 8 9 10 11 12 13 14
do
(
~/meme/bin/tomtom \
  -no-ssc -verbosity 1 -min-overlap 5 -dist pearson -evalue -thresh 10.0 \
  code/all_denovo_pwms_V3/all_denovo_pwms_V3.meme${batch} \
  motif_databases/MOUSE/HOCOMOCOv11_full_MOUSE_mono_meme_format.meme \
  motif_databases/HUMAN/HOCOMOCOv11_full_HUMAN_mono_meme_format.meme \
  -text > all_denovo_PWM_tomtom_v3/all_denovo_PWM_tomtom_v3_${batch}.tsv
) &
done

# random (null) results

# ~/meme/bin/tomtom -no-ssc -oc random_PWM_tomtom -verbosity 1 -min-overlap 5 -dist pearson -evalue -thresh 10.0 code/random_pwms.meme motif_databases/MOUSE/HOCOMOCOv11_full_MOUSE_mono_meme_format.meme motif_databases/HUMAN/HOCOMOCOv11_full_HUMAN_mono_meme_format.meme
```


# Load tomtom results

```{r, fig.width=4}
#load results

results_all <- readRDS("../data/motifs/motifFinder_denovo_results_all_clean_V3.rds")

tomtom_summary <- rbindlist(lapply(list.files(path = "../data/motifs/all_denovo_PWM_tomtom_v3/", full.names = T),
                                   function(x) fread(cmd=paste("grep -v '#'",x), check.names = T)))

```


```{r}

# Calculate and add Information Content values

IC_values <- unlist(lapply(1:9000, information_content))

textPWM <- function(x){
  if(!is.null(results_all[[x]])){
    return(pwm2text(results_all[[x]]$scoremat))
  }else{
    return(NA)
  }
}

textPWMs <- sapply(1:9000, textPWM)

# need to convert NULL to NA via char
maxiters <- as.integer(as.character(lapply(results_all,function(x) nrow(x$alphas))))

IC_values <- data.table(IC_values, maxiters, Query_ID=names(results_all), textual_PWM=textPWMs)
setkey(IC_values,Query_ID)
setkey(tomtom_summary,Query_ID)

tomtom_summary <- merge(tomtom_summary, IC_values, all.x=T)

# extract / format transcription factor name
tomtom_summary[,Target.Name := sapply(strsplit(Target_ID,"_"), function(x) x[[1]][[1]])]
#tomtom_summary_random[,Target.Name := sapply(strsplit(Target_ID,"_"), function(x) x[[1]][[1]])]

# extract / format component name
tomtom_summary[,Component := sapply(strsplit(Query_ID,"\\."), function(x) x[[1]][[1]])]
#tomtom_summary_random[,Component := sapply(strsplit(Query_ID,"\\."), function(x) x[[1]][[1]])]

# calculate GC statistics from consensus
tomtom_summary[,GC_count := nchar(gsub("[AT]","",Query_consensus))]
tomtom_summary[,AT_count := nchar(gsub("[GC]","",Query_consensus))]
tomtom_summary[,GC_percent := (GC_count)/(AT_count+GC_count)]

# Extract Iteration ID
tomtom_summary[,Iteration := sapply(strsplit(Query_ID,"\\."), function(x) as.numeric(gsub("I","",x[[2]][[1]])))]

tomtom_summary[Iteration<=9*3, region:="Right"]
tomtom_summary[Iteration>9*3 & Iteration<=9*6, region:="Left"]
tomtom_summary[Iteration>9*6, region:="Centered"]

tomtom_summary[,IC_perbase := IC_values/nchar(Query_consensus)]

# for each denovo motif, use best tomtom match
tomtom_summary_byquery <- tomtom_summary[,.SD[which.min(E.value)],by=Query_ID]

fwrite(tomtom_summary_byquery, "../data/motifs/tomtom_summary_byQuery.tsv")
```


# GC

```{r}
hist(tomtom_summary_byquery$GC_percent, breaks=200)

ggplot(tomtom_summary_byquery, aes(nchar(Query_consensus), GC_percent)) + geom_point(alpha=0.005) + scale_x_log10()
```

# Most frequent matches
```{r}

tomtom_summary_byquery[,.N,by=Target.Name][order(-N)][1:40]

```

# Are 'unconverged' motifs worse?

```{r}

ggplot(tomtom_summary_byquery, aes(IC_values, colour=maxiters==1000)) + geom_density()

ggplot(tomtom_summary_byquery, aes(-log10(p.value), colour=maxiters==1000)) + geom_density()

```


# for each most common target plot match
```{r}
for(match in tomtom_summary_byquery[,.N,by=Target.Name][order(-N)][1:50]$Target.Name){
  print(plot_motif_matches(tomtom_summary_byquery[Target.Name==match][order(p.value)][1]))
}
```

# which motifs have NO TOMTOM match
```{r}
for(match in names(which(sapply(results_all, is.null)==F))[!names(which(sapply(results_all, is.null)==F)) %in% unique(tomtom_summary_byquery$Query_ID)]){
  print(ggseqlogo(get_PWM(results_all[[match]])))
}
```


# Does search region affect motif quality
```{r}
ggplot(tomtom_summary_byquery[,.("mean_IC"=mean(IC_values)),by=c("Iteration","region")], aes(Iteration, mean_IC, colour=region)) + geom_point()
ggplot(tomtom_summary_byquery[,.("mean_pvalue"=mean(-log10(p.value))),by=c("Iteration","region")], aes(Iteration, mean_pvalue, colour=region)) + geom_point()
```





# Look for novel motifs

Look for motifs with high information content but low match to currently known motifs

Most high IC are GC rich - probably SP1 etc.

```{r, fig.width=6}

ggplot(tomtom_summary_byquery, aes(E.value, IC_perbase, colour=GC_percent)) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom") +
  scale_colour_gradientn(colours = viridis(100, direction = -1))

ggplot(tomtom_summary_byquery, aes(E.value, IC_perbase, colour=grepl("^SP[0-9]",Target.Name))) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom")

ggplot(tomtom_summary_byquery, aes(E.value, IC_perbase, colour=grepl("ATF[0-9]",Target.Name))) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom")

# Stra8? motif
ggplot(tomtom_summary_byquery, aes(E.value, IC_perbase, colour=grepl("A?GG[AT][AG][AG]?|[TC]?[TC][TA]CCT?",Query_consensus))) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom")


ggplot(tomtom_summary_byquery, aes(E.value, IC_perbase, colour=grepl("CREB1",Target.Name))) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom")

ggplot(tomtom_summary_byquery, aes(E.value, IC_perbase, colour=IC_perbase>0.8 & E.value>1e-4)) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom")
```

##  C26

```{r}

print(ggseqlogo(get_PWM(results_all[paste0("C26N.I",69)][[1]],T)) + ggtitle(paste(i,"T")))
for(i in c(65,68,70)){
  print(ggseqlogo(get_PWM(results_all[paste0("C26N.I",i)][[1]])) + ggtitle(i))
}

```

```{r}
for(i in c(14,22)){
  print(ggseqlogo(get_PWM(results_all[paste0("C26N.I",i)][[1]])) + ggtitle(i))
}
print(ggseqlogo(get_PWM(results_all[paste0("C26N.I",67)][[1]],T)) + ggtitle(paste(i,"T")))

```

## 3 C's

```{r}

# CXXCXXC
for(i in c(20,15,80,1,6)){
  print(ggseqlogo(get_PWM(results_all[paste0("C2P.I",i)][[1]])))
}

# CXXCXXC
for(i in c(13,1,27,79)){
  print(ggseqlogo(get_PWM(results_all[paste0("C39P.I",i)][[1]],T)))
}

# CXXCXXC
for(i in c(5,62,63,52,59,79,8,70,74,56)){
  print(ggseqlogo(get_PWM(results_all[paste0("C48P.I",i)][[1]],T)))
}

```


```{r}
tomtom_summary_byquery[IC_perbase>0.5 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)][,.N,by=Component][order(-N)]
tomtom_summary_byquery[IC_perbase>0.5 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)][,.N,by=Target.Name][order(-N)]
tomtom_summary_byquery[IC_perbase>0.5 & E.value>1e-3 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)][,.N,by=Target.Name][order(-N)]

tomtom_summary_byquery[IC_perbase>0.7 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)][!grepl("A?GG[AT][AG][AG]?|[TC]?[TC][TA]CCT?",Query_consensus) & !grepl("ATF[0-9]|CREB1",Target.Name)][order(Component)]


# C28P = ATF2&7
# C41P = CREB1

# C48P - CXXCXXC|GXXGXXG

for(i in 1:20){
  print(plot_motif_matches(tomtom_summary_byquery[IC_perbase>0.5 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)][Component=="C48P"][i])) # 41-45 c32, C39 47-49  
}

# CXXCXXC again, some TTT
for(i in 1:19){
  print(plot_motif_matches(tomtom_summary_byquery[IC_perbase>0.5 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)][Component=="C39P"][i])) # 41-45 c32, C39 47-49  
}


for(i in 1:17){
  print(plot_motif_matches(tomtom_summary_byquery[IC_perbase>0.5 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)][Component=="C2P"][i])) # 41-45 c32, C39 47-49  
}


for(i in 1:12){
  print(plot_motif_matches(tomtom_summary_byquery[IC_perbase>0.5 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)][Component=="C50P"][i])) # 41-45 c32, C39 47-49  
}
# TXXXXXT

for(i in 1:14){
  print(plot_motif_matches(tomtom_summary_byquery[IC_perbase>0.5 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)][Component=="C33P"][i])) # 41-45 c32, C39 47-49  
}

for(i in 1:16){
  print(plot_motif_matches(tomtom_summary_byquery[IC_perbase>0.5 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)][Component=="C31P"][i])) # 41-45 c32, C39 47-49  
}


# AXCXXC
for(i in c(36,83,4)){ # 26,61, opposite side component, RFX
  plot_motif_matches(tomtom_summary_byquery[Query_ID==paste0("C33P.I",i)])
}




for(i in 1:5){
  print(plot_motif_matches(tomtom_summary_byquery[IC_perbase>0.7 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)][!grepl("A?GG[AT][AG][AG]?|[TC]?[TC][TA]CCT?",Query_consensus) & !grepl("ATF[0-9]|CREB1",Target.Name)][Target.Name=="TBP"][i])) # 41-45 c32, C39 47-49  
}

for(i in 1:4){
  print(plot_motif_matches(tomtom_summary_byquery[IC_perbase>0.7 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)][!grepl("A?GG[AT][AG][AG]?|[TC]?[TC][TA]CCT?",Query_consensus) & !grepl("ATF[0-9]|CREB1",Target.Name)][Target.Name=="OSR2"][i])) # 41-45 c32, C39 47-49  
}

for(i in 1:4){
  print(plot_motif_matches(tomtom_summary_byquery[IC_perbase>0.7 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)][!grepl("A?GG[AT][AG][AG]?|[TC]?[TC][TA]CCT?",Query_consensus) & !grepl("ATF[0-9]|CREB1",Target.Name)][Target.Name=="BHA15"][i])) # 41-45 c32, C39 47-49  
}

# GTGAGT|ACTCAC - MAFG like, 14N (sc) & 3N (lympho)

for(i in 1:5){
  print(plot_motif_matches(tomtom_summary_byquery[IC_perbase>0.7 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)][Component=="C14N"][i])) # 41-45 c32, C39 47-49  
}


for(i in 1:20){
  print(plot_motif_matches(tomtom_summary_byquery[IC_perbase>0.5 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)][Target.Name=="TAF1"][i])) # 41-45 c32, C39 47-49  
}


for(i in 1:6){
  print(plot_motif_matches(tomtom_summary_byquery[IC_perbase>0.7 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)][Target.Name=="ZN770"][i])) # 41-45 c32, C39 47-49  
}

for(i in 1:5){
  print(plot_motif_matches(tomtom_summary_byquery[IC_perbase>0.7 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)][Target.Name=="ZEB1"][i])) # 41-45 c32, C39 47-49  
}

for(i in 1:12){
  print(plot_motif_matches(tomtom_summary_byquery[IC_perbase>0.7 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)][Target.Name=="ETV7"][i])) # 41-45 c32, C39 47-49  
}

for(i in 1){
  print(plot_motif_matches(tomtom_summary_byquery[IC_perbase>0.7 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)][Target.Name=="ETV6"][i])) # 41-45 c32, C39 47-49  
}
# 21N - 2

tomtom_summary_byquery[IC_perbase>0.7 & E.value>1e-1]
tomtom_summary_byquery[IC_perbase>0.7 & E.value>1e-1 & grepl("A?GG[AT][AG][AG]?|[TC]?[TC][TA]CCT?", Query_consensus)]

tomtom_summary_byquery[IC_perbase>0.7 & E.value>1e-1 & grepl("CAGCAG", Query_consensus)] #Prdm9like

tomtom_summary_byquery[IC_perbase>0.7 & E.value>1e-1 & Component=="C5N"]

tomtom_summary_byquery[IC_perbase>0.5 & grepl("TGAGG.[GA]{2}|[CT]{2}.CCTCA", Query_consensus)]



for(i in c(1:90)){
  if(!is.null(results_all[paste0("C44N.I",i)][[1]])){
    print(ggseqlogo(get_PWM(results_all[paste0("C44N.I",i)][[1]]))+ ggtitle(paste0("C44N.I",i)))  
  }
}
```

# Export for regprob finding

```{r}
plot_motif_matches(tomtom_summary_byquery[IC_perbase>0.5 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)  & IC_values>5][,.SD[which.max(IC_perbase)],by=Target_ID][4])

test_queries <- tomtom_summary_byquery[IC_perbase>0.5 & E.value>1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)  & IC_values>5][,.SD[which.max(IC_perbase)],by=Target_ID]$Query_ID

# V2
test_queries <- tomtom_summary_byquery[IC_perbase>0.5 & E.value>1e-5 & E.value<1e-1 & !grepl("ATF[0-9]|CREB1",Target.Name) & (GC_count+AT_count>4)  & IC_values>5][,.SD[which.max(IC_perbase)],by=Target_ID]$Query_ID


ggplot(tomtom_summary_byquery, aes(E.value, IC_perbase, colour=Query_ID %in% test_queries)) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom")

dput(test_queries)
```


#  31

```{r}
# TXXTXXT
for(i in c(18,13,11)){
  print(plot_motif_matches(tomtom_summary_byquery[Query_ID==paste0("C31P.I",i)]))
}
```

```{r}
for(i in c(6,3,20,26,23,59)){
  print(plot_motif_matches(tomtom_summary_byquery[Query_ID==paste0("C31P.I",i)]))
}
```

```{r}
plot_motif_matches(tomtom_summary_byquery[Query_ID=="C31P.I19"])
plot_motif_matches(tomtom_summary_byquery[Query_ID=="C31P.I6"])
plot_motif_matches(tomtom_summary_byquery[Query_ID=="C31P.I20"])
plot_motif_matches(tomtom_summary_byquery[Query_ID=="C31P.I26"])
plot_motif_matches(tomtom_summary_byquery[Query_ID=="C31P.I59"])
```


```{r}
# GGXXXXGG

for(i in c(19,25,17,24,10,59,12,22,2)){
  print(plot_motif_matches(tomtom_summary_byquery[Query_ID==paste0("C31P.I",i)]))
}

```


If we look at motifs with GC of between 10% and 90%, there's a lump of motifs with relatively high IC and low E value.

```{r, fig.width=6}

ggplot(tomtom_summary_byquery[GC_percent<0.9 & GC_percent>0.1], aes(E.value, IC_values, colour=GC_percent)) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom") +
  scale_colour_gradientn(colours = viridis(100))

ggplot(tomtom_summary_byquery[GC_percent<0.9 & GC_percent>0.1], aes(E.value, IC_values, colour=GC_percent<0.8 & GC_percent>0.2 & E.value>1e-4 & IC_values>10)) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom")
```

They mostly match (badly) with ATF1 and are from the late components 17, 18, 36, 41, and 34.

```{r, fig.width=6}

tomtom_summary_byquery[GC_percent<0.8 & GC_percent>0.2][E.value>1e-4 & IC_values>10][,.N,by=Target.Name][order(-N)]

tomtom_summary_byquery[E.value>1e-2 & IC_perbase>0.5][,.N,by=Target.Name][order(-N)]

tomtom_summary_byquery[GC_percent<0.8 & GC_percent>0.2][E.value>1e-4 & IC_values>10][,.N,by=Component][order(-N)][1:10]

ggplot(tomtom_summary_byquery, aes(E.value, IC_values, colour=grepl("ATF",Target.Name))) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom")

ggplot(tomtom_summary_byquery, aes(E.value, IC_values, colour=grepl("TGA[ACGT]GTCACAA?|T?TGTGAC[ACGT]TCA",Query_consensus))) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom")

ggplot(tomtom_summary_byquery, aes(E.value, IC_values, colour=grepl("^SP",Target.Name))) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom")

ggplot(tomtom_summary_byquery, aes(E.value, IC_perbase, colour=grepl("^RFX",Target.Name))) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom") + scale_color_manual(values=c("white","black"))


ggplot(tomtom_summary_byquery, aes(E.value, IC_values, colour=grepl("^TAF1",Target.Name))) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom") + scale_color_manual(values=c("white","black"))


ggplot(tomtom_summary_byquery, aes(E.value, IC_perbase, colour=grepl("GG[ACT]GG[ACT]GG?|CC[AGT]CC[AGT]CC",Query_consensus))) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom")


```

Looking at a few examples they all look pretty much the same, part of an ATF1 motif, plus CACAA. The ATF1 motif itself is quite weak also.

Perhaps the most interesting part of these motifs, is that the CpG part of them is much reduced or absent compared to the currently accepted Crem motif.

```{r, fig.width=6}

set.seed(42)

cowplot::plot_grid(plotlist =  list(
plot_motif_matches(tomtom_summary_byquery[E.value>1e-2 & & IC_perbase>0.5][sample(1:1638, 1)], yaxis = T)[[1]],
plot_motif_matches(tomtom_summary_byquery[GC_percent<0.8 & GC_percent>0.2][E.value>1e-4 & IC_values>10][sample(1:172, 1)], yaxis = T)[[1]],
plot_motif_matches(tomtom_summary_byquery[GC_percent<0.8 & GC_percent>0.2][E.value>1e-4 & IC_values>10][sample(1:172, 1)], yaxis = T)[[1]],
plot_motif_matches(tomtom_summary_byquery[GC_percent<0.8 & GC_percent>0.2][E.value>1e-4 & IC_values>10][sample(1:172, 1)], yaxis = T)[[1]])
)

plot_motif_matches(tomtom_summary_byquery[Query_ID=="C18P.I53"], yaxis = T)

# export_PWM(results_all[["C18P.I53"]],"C18P.I53", file="C18P.I53.motif")

```

Maybe we're biasing towards small amount of CpG by conditioning on (poor) match with ATF1, try using generic consensus motif pattern instead: there are some others not matching ATF1, there are 14, and they are in the top 15 motifs ordered by E.value. They (poorly) match ATF7, RXRB, and CREB1.

```{r, fig.width=6}

ggplot(tomtom_summary_byquery, aes(E.value, IC_values, colour=IC_values>10 & E.value>1e-4 & Target.Name!="ATF1" & grepl("TGTGA|TCACA", Query_consensus))) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom")

cowplot::plot_grid(plotlist =  list(
plot_motif_matches(tomtom_summary_byquery[E.value>1e-4 & IC_values>10 & grepl("TGTGA|TCACA", Query_consensus)][order(-E.value)][1], yaxis = T)[[1]],
plot_motif_matches(tomtom_summary_byquery[E.value>1e-4 & IC_values>10 & grepl("TGTGA|TCACA", Query_consensus)][order(-E.value)][2], yaxis = T)[[1]],
plot_motif_matches(tomtom_summary_byquery[E.value>1e-4 & IC_values>10 & grepl("TGTGA|TCACA", Query_consensus)][order(-E.value)][4], yaxis = T)[[1]],
plot_motif_matches(tomtom_summary_byquery[E.value>1e-4 & IC_values>10 & grepl("TGTGA|TCACA", Query_consensus)][order(-E.value)][5], yaxis = T)[[1]])
)

```

```{r}
ggplot(tomtom_summary_byquery, aes(E.value, IC_values, colour=Query_ID=="C5N.I82")) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom")

ggplot(tomtom_summary_byquery, aes(E.value, IC_values, colour=region)) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom")

ggplot(tomtom_summary_byquery, aes(E.value, IC_values, colour=region)) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom") + facet_wrap(~region) + ylim(NA,20)

ggplot(tomtom_summary_byquery, aes(E.value, IC_perbase, colour=region)) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom") + facet_wrap(~region)

ggplot(tomtom_summary_byquery, aes(E.value, IC_perbase, colour=Iteration)) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom") + scale_color_viridis()

tomtom_summary_byquery[region=="Right" & IC_perbase>0.75][,.N,by=Component][order(-N)]

tomtom_summary_byquery[region=="Right" & IC_perbase>0.6][Component=="C44N"]

plot_motif_matches(tomtom_summary_byquery[region=="Right" & IC_perbase>0.7][Component=="C15N"][3])

plot_motif_matches(tomtom_summary_byquery[region=="Right" & IC_perbase>0.75][Component=="C5N"][1])

tomtom_summary_byquery[region=="Right"][order(-IC_perbase)]

ggplot(tomtom_summary_byquery, aes(E.value, IC_perbase, colour=grepl("TGAGG|CCTCA", Query_consensus))) +
  geom_point(size=0.2) +
  scale_x_log10() +
  theme(legend.position = "bottom")



# highest IC, match to Stra8 consensus
tomtom_summary_byquery[grepl("TGAGG.[GA]{2}|[CT]{2}.CCTCA", Query_consensus)][order(-IC_perbase)]
plot_motif_matches(tomtom_summary_byquery[grepl("TGAGG.[GA]{2}|[CT]{2}.CCTCA", Query_consensus)][order(-IC_perbase)][1])

# highest IC, match to CremT
plot_motif_matches(tomtom_summary_byquery[grepl("TGA[ACGT]GTCACAA|TTGTGAC[ACGT]TCA",Query_consensus)][order(-IC_perbase)][1])


for(i in 1:20){
  print(plot_motif_matches(tomtom_summary_byquery[region=="Right"][order(-IC_perbase)][i]))
}
```


## GGCGGC

Anything else after excluding all matched to the ATF1 like motif? Mostly long GGCGGCGG like motifs. TF motif or some kind of CpG island?

```{r, fig.width=6}

ggplot(tomtom_summary_byquery, aes(E.value, IC_perbase, colour=IC_perbase>0.8 & E.value>1e-4 & !grepl("TGTGA|TCACA", Query_consensus))) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom")

ggplot(tomtom_summary_byquery, aes(E.value, IC_values, colour=IC_values>10 & E.value>1e-4 & !grepl("TGTGA|TCACA", Query_consensus))) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom")

tomtom_summary_byquery[IC_values>10 & E.value>1e-4 & !grepl("TGTGA|TCACA", Query_consensus)]
plot_motif_matches(tomtom_summary_byquery[IC_values>10 & E.value>1e-4 & !grepl("TGTGA|TCACA", Query_consensus) & Component!="C19N"][order(-E.value)][1], yaxis = T)
plot_motif_matches(tomtom_summary_byquery[IC_values>10 & E.value>1e-4 & !grepl("TGTGA|TCACA", Query.consensus) & Component!="C19N"][order(-E.value)][6], yaxis = T)
plot_motif_matches(tomtom_summary_byquery[IC_values>10 & E.value>1e-4 & !grepl("TGTGA|TCACA", Query.consensus) & Component!="C19N"][order(-E.value)][7], yaxis = T)
plot_motif_matches(tomtom_summary_byquery[IC_values>10 & E.value>1e-4 & !grepl("TGTGA|TCACA", Query.consensus) & Component!="C19N"][order(-E.value)][8], yaxis = T)

```

These motifs mostly match (poorly) with TAF1 & SP1,2,3. Diverse components, a bit batchy.

```{r, fig.width=6}

ggplot(tomtom_summary_byquery, aes(E.value, IC_values, colour=grepl("CCGCC|GGCGG", Query.consensus))) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom")

ggplot(tomtom_summary_byquery, aes(E.value, IC_values, colour=grepl("GCCGCCG|CGGCGGC", Query.consensus))) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom")

tomtom_summary_byquery[grepl("GCCGCCG|CGGCGGC", Query.consensus)][E.value>1e-4 & IC_values>10][,.N,by=Target.Name][order(-N)][1:10]

tomtom_summary_byquery[grepl("GCCGCCG|CGGCGGC", Query.consensus)][E.value>1e-4 & IC_values>10][,.N,by=Component][order(-N)][1:10]

ggplot(tomtom_summary_byquery, aes(E.value, IC_values, colour=Target.Name %in% c("TAF1"))) +
  geom_point(size=0.1) +
  scale_x_log10() +
  theme(legend.position = "bottom")

plot_motif_matches(tomtom_summary_byquery[Target.Name %in% c("TAF1")][order(-IC_values)][1], yaxis = T)

tomtom_summary_byquery[grepl("GCCGCCG|CGGCGGC", Query.consensus) & nchar(Query.consensus)<20][order(-IC_values)][1:20]

cowplot::plot_grid(plotlist =  list(
plot_motif_matches(tomtom_summary_byquery[grepl("GCCGCCG|CGGCGGC", Query.consensus) & nchar(Query.consensus)<20][order(-IC_values)][1], yaxis = T)[[1]],
plot_motif_matches(tomtom_summary_byquery[grepl("GCCGCCG|CGGCGGC", Query.consensus) & nchar(Query.consensus)<20][order(-IC_values)][2], yaxis = T)[[1]],
plot_motif_matches(tomtom_summary_byquery[grepl("GCCGCCG|CGGCGGC", Query.consensus) & nchar(Query.consensus)<20][order(-IC_values)][3], yaxis = T)[[1]],
plot_motif_matches(tomtom_summary_byquery[grepl("GCCGCCG|CGGCGGC", Query.consensus) & nchar(Query.consensus)<20][order(-IC_values)][4], yaxis = T)[[1]])
)

```

More likely to match mouse motif than human

```{r}
ggplot(tomtom_summary, aes(p.value)) + geom_density(aes(colour=grepl("_MOUSE",Target_ID))) + scale_x_log10()  + ylim(0,0.1)
tomtom_summary[,mean(log10(p.value)),by=grepl("MOUSE",Target_ID)]
tomtom_summary_byquery[,.N,by=grepl("MOUSE",Target_ID)]
```


# Grouped
For each motif group, plot the denovo motifs

```{r, fig.width=6}
# Which motifs have best matches
# best_denovo <- tomtom_summary[,.SD[which.min(p.value)],by=Target_ID][order(E.value)][1:400]
# tomtom_summary_byquery[,min(E.value),by=Target.Name][order(V1)][1:20]


# for each TF, select the best matching denovo motif, for those with min q of 0.001
#q.value<0.001
motif_matches <- tomtom_summary[E.value<1e-3][, .SD[which.min(p.value)], by = Target.Name][order(E.value)]
saveRDS(motif_matches, "../data/motifs/motif_matches_v3.rds")
fwrite(motif_matches, "../data/motifs/tomtom_summary_byTarget.tsv")

print(paste(nrow(motif_matches), "different matches"))
# 123 (126v1) matches
print(paste(length(unique(motif_matches$Query_ID)),"different motifs"))
# from 101 (92v1) motifs

pdf("../results/known_motifs_v3.pdf")
for(i in 1:nrow(motif_matches)){
  print(plot_motif_matches(motif_matches[order(q.value)][i]))
}
dev.off()

# load motif match groupings
motif_groups <- read.table("../data/motifs/motif_results_groups.csv", fill=T, sep = ",", stringsAsFactors=F)
motif_groups$group <- c(1:(nrow(motif_groups)))
motif_groups <- data.table(melt(motif_groups, id.vars="group"))[value!=""]

print(paste("Groups:",paste(unique(motif_groups$group), collapse = " ")))

# check all the matches are present
sum(!motif_groups$value %in% motif_matches$Target.Name)
sum(!motif_matches$Target.Name %in% motif_groups$value)

motif_groups$value[!motif_groups$value %in% motif_matches$Target.Name]
motif_matches$Target.Name[!motif_matches$Target.Name %in% motif_groups$value]

# for each group of motifs plot the matches

pdf("../results/Denovo_motifs_grouped_V3B.pdf", width = 9)
for(class in 1:max(unique(motif_groups$group), na.rm = T)){
  print(class)
  tmp <- motif_groups[group==class]$value
  if(nrow(motif_matches[Target.Name %in% tmp])>0){
    print(plot_grid(plotlist = plot_motif_matches(motif_matches[Target.Name %in% tmp], titles = T)))
  }else{
    print(paste("NO",tmp))
  }
  
}
dev.off()

```

see results/Denovo_motifs_grouped.pdf for plots of grouped motifs


In which components are the motifs found

```{r, fig.width=6}

#best_matches <- c("THA11","SP2","NFYB","RFX2","NRF1","GABPA","SPIB","MYB","ATF2","ETS1","ETV5","TBP","CREB1","THAP1","TYY1")

ordering <- c(rbind(paste0("C",component_labels,"P"),paste0("C",component_labels,"N")))

#lapply(ordering, function(x) tomtom_summary_byquery[Component==x][Target.Name %in% best_matches, min(E.value), by=Target.Name])

# For each TF - component combo, get min E.Value

motif_which_component <- tomtom_summary[Target.Name %in% motif_groups[!is.na(group)]$value, min(E.value), by=.(Component,Target.Name)][order(Target.Name)]

motif_which_component <- dcast(motif_which_component, formula= Component ~ Target.Name, value.var="V1")

motif_which_component_m <- as.matrix(motif_which_component[,-"Component", with=F])
rownames(motif_which_component_m) <- motif_which_component$Component

ordering2 <- ordering[ordering %in% motif_which_component$Component][1:75] #46

motif_which_component_m[is.na(motif_which_component_m)] <- 12

motif_which_component_m <- -log(motif_which_component_m[ordering2,],10)

aheatmap(motif_which_component_m, Rowv = NA, cexRow = 1, scale = "col", hclustfun = "ward.D2")

motif_which_component_m[motif_which_component_m>10] <- 10

aheatmap(motif_which_component_m, Rowv = NA, cexRow = 1, hclustfun = "ward.D2")

```

## Stra8 Motif

from personal communication with Kojima

```{r}
stra8_motif <- fread("../data/previous_studies/Kojima/stra8_motif1.motif", skip = 1)
stra8_motif <- t(as.matrix(stra8_motif))
rownames(stra8_motif) <- c("A","C","G","T")
ggseqlogo(as.matrix(stra8_motif))

export_PWM(as.matrix(stra8_motif), name = "Stra8_Kojima_Elife",file = "../data/previous_studies/Kojima/Stra8_Kojima_Elife.meme")
```

# Plot best Motif per group

```{r, fig.asp=0.5}
# print best matches in grid

best_matches <- c("SP2", "NFYA", "ZN143", "RFX6", "NRF1", "GABPA", "SPI1", "ATF1", "ETV5", "TYY1", "RARA", "TBP", "MYBA","TAF1")

best_denovo <- data.table()
for (i in best_matches){
  best_denovo <- rbind(best_denovo, tomtom_summary[grepl(i,Target_ID)][order(p.value)][1]) #[!grepl("D$",Target_ID)]
}

# Add novel
best_denovo <- rbind(best_denovo, tomtom_summary_byquery[Query_ID %in% c("C5N.I57","C17P.I37")]) #"C7N.I82","C18P.I53" (TAF1 & ATF1), C5.25 26, 15
saveRDS(best_denovo, "../data/best_denovo_V3.rds")

best_denovo <- readRDS("../data/best_denovo_V3.rds")

# make Stra8 match new motif (not yet in HocoMoco)
best_denovo[Query_ID=="C5N.I57"]$Target_ID <- "LOCAL../data/previous_studies/Kojima/Stra8_Kojima_Elife.meme"
best_denovo[Query_ID=="C5N.I57"]$Optimal_offset <- -1
best_denovo[Query_ID=="C5N.I57"]$Orientation  <- "-"

matched_pwms <- plot_motif_matches(best_denovo,
                                   titles=c("Sp2",
                                            "Nfya",
                                            "Zfp143",
                                            "Rfx2 (Rfx6)",
                                            "Nrf1", #5
                                            "Gabpa",
                                            "Spi1",
                                            "Crem (Atf1)",
                                            "Etv5",
                                            "Zfp42 (Yy1)", #10
                                            "Rara",
                                            "Tbp",
                                            "Mybl1",
                                            "Taf1",
                                            "Crem-t (Atf1)", #15
                                            "Stra8*"))

# quick plot to check
pdf("../results/motifs/Motis_V3.pdf")
plot_grid(plotlist=matched_pwms, ncol=4)
dev.off()


# Make final plot with icon annotations and in order

x_pos = 0.65
y_pos = 0.82
wh <- 0.16
icons <- c("NW","TL","TI","TI","IW","NW","MW","TI","IW","TW","IW","TL","T2I","NW","TI","TI")
tmp <- list()

for(i in seq_along(matched_pwms)){
  tmp[[i]] <- ggdraw() +
    draw_plot(matched_pwms[[i]], 0, 0, 1, 1) +
    draw_image(paste0("../results/icons/",icons[i],".png"), x=x_pos, y=y_pos, width=wh, height=wh)
}

# reorder to match heatmap
# c(1,9,16,3,5,6,2,10,14,11,8,4,13,12,7,15)
tmp = tmp[c(1,14,6,5,9,10,2,3,16,13,4,8,11,12,7,15)] 

saveRDS(tmp,"../data/plots/denovo_motifs.rds")
tmp <- readRDS("../data/plots/denovo_motifs.rds")

library(cowplot)

pdf(height=5, width=10,file="../results/motifs/motifs_fig_A.pdf")
plot_grid(plotlist=tmp, ncol=4)
dev.off()

```


# Calculate regprobs using found motifs pwms

```{r}
results_all <- readRDS("motifFinder_denovo_results_all_clean_V3.rds")

motif_matches <- readRDS("../data/motifs/motif_matches_v3.rds")

SDAresults <- load_results(results_folder = "../data/conradV3/conradV3_sda_1/", data_path = "../data/conradV3/")
rownames(SDAresults$loadings[[1]]) <- paste0("V",1:50)


tss_seqs <- seqinr::read.fasta("../data/motifs/promoter_sequences/all_genes_tss_1000.fasta", forceDNAtolower = FALSE, as.string = TRUE)
tss_seqs_s <- toupper(as.character(tss_seqs))
names(tss_seqs_s) <- gsub("\\(-\\)|\\(\\+\\)","",names(tss_seqs))

tss_seqs_s <- substr(tss_seqs_s,500,1500)

tss_seqs_s <- tss_seqs_s[names(tss_seqs_s) %in% colnames(SDAresults$loadings[[1]])]


pwms <- lapply(results_all[c(motif_matches$Query_ID,"C5N.I57","C17P.I37")], get_PWM)

names(pwms) <- c(motif_matches$Target.Name,"STRA8","CREMT")

library(parallel)
system.time(results_custom <- mclapply(pwms, function(x) find_motifs_parallel(x, custom=T), mc.cores = 16, mc.silent=FALSE, mc.preschedule = FALSE))

saveRDS(results_custom, "../data/motifs/MF_results_custom_V3b.rds")

```

# Regprobs for C2N

```{r}

queryIDs <- strsplit(paste0("C2N.I",1:90,collapse = " ")," ")[[1]]
queryIDs <- queryIDs[-which(sapply(results_all[queryIDs], is.null))]

pwms <- lapply(results_all[queryIDs], get_PWM)



library(parallel)
system.time(results_C2N <- mclapply(pwms, function(x) find_motifs_parallel(x, custom=T), mc.cores = 16, mc.silent=FALSE, mc.preschedule = FALSE))

saveRDS(results_C2N, "../data/motifs/MF_results_C2N_V3.rds")


####

queryIDs <- strsplit(paste0("C5N.I",1:90,collapse = " ")," ")[[1]]
queryIDs <- queryIDs[-which(sapply(results_all[queryIDs], is.null))]

pwms <- lapply(results_all[queryIDs], get_PWM)

library(parallel)
system.time(results_C5N <- mclapply(pwms, function(x) find_motifs_parallel(x, custom=T), mc.cores = 16, mc.silent=FALSE, mc.preschedule = FALSE))

saveRDS(results_C5N, "../data/motifs/MF_results_C5N_V3.rds")


###

queryIDs <- strsplit(paste0("C11P.I",1:90,collapse = " ")," ")[[1]]
queryIDs <- queryIDs[-which(sapply(results_all[queryIDs], is.null))]

pwms <- lapply(results_all[queryIDs], get_PWM)

library(parallel)
system.time(results_C11P <- mclapply(pwms, function(x) find_motifs_parallel(x, custom=T), mc.cores = 16, mc.silent=FALSE, mc.preschedule = FALSE))

saveRDS(results_C11P, "../data/motifs/MF_results_C11P_V3.rds")

####

queryIDs <- strsplit(paste0("C3P.I",1:90,collapse = " ")," ")[[1]]
queryIDs <- queryIDs[-which(sapply(results_all[queryIDs], is.null))]

pwms <- lapply(results_all[queryIDs], get_PWM)

library(parallel)
system.time(results_C3P <- mclapply(pwms, function(x) find_motifs_parallel(x, custom=T), mc.cores = 16, mc.silent=FALSE, mc.preschedule = FALSE))

saveRDS(results_C3P, "../data/motifs/MF_results_C3P_V3.rds")

####


pwms <- lapply(results_all[test_queries], get_PWM)

system.time(results_novel <- mclapply(pwms, function(x) find_motifs_parallel(x, custom=T), mc.cores = 16, mc.silent=FALSE, mc.preschedule = FALSE))

saveRDS(results_novel, "../data/motifs/MF_results_novel_B.rds")

#saveRDS(results_novel, "../data/motifs/MF_results_novel_A.rds")

#    user    system   elapsed
# 25261.987 11709.344  2385.375
```


```{r, fig.width=10}
results_custom <- readRDS("../data/motifs/MF_results_custom_V3.rds")
tmp <- names(results_custom)
results_custom <- sapply(1:length(results_custom), function(x) results_custom[[x]]$regprob)
colnames(results_custom) <- tmp

library(ComplexHeatmap)
Heatmap(cor(results_custom))
Heatmap(cor(results_custom, method = "spearman"))
```

# Export Stra8

To run FIMO to predict targets & assess specificity

```{r}
plot_motif_matches(tomtom_summary_byquery[Query_ID=="C5N.I26"])

export_PWM(get_PWM(results_all["C5N.I26"][[1]]), "C5N.I26", "C5N.I26.meme")

fimo_stra8 <- fread("~/Downloads/fimo.tsv")
fimo_stra8[,.N,by=sequence_name][order(-N)]

fimo_stra8[,.N,by=nchar(sequence_name)>5][order(-N)]
```
