---
title: "MotifFinder Overview"
output: 
  html_notebook: 
    code_folding: hide
    results: hold
    fig_show: hold
    toc: yes
    toc_float: yes
    number_sections: yes
    theme: spacelab
---



# CpG Count as a motif

```{r, fig.width=6}

tss_seqs <- seqinr::read.fasta("../data/motifs/promoter_sequences/all_genes_tss_500.fasta", forceDNAtolower = FALSE, as.string = TRUE)
tss_seqs_s <- toupper(as.character(tss_seqs))
names(tss_seqs_s) <- gsub("\\(-\\)|\\(\\+\\)","",names(tss_seqs))

CG_count <- sapply(tss_seqs_s,function(x) stringr::str_count(x,"CG"))
CG_locations <- sapply(tss_seqs_s,function(x) stringr::str_locate_all(x,"CG")[[1]][,1])

hist(CG_count, breaks=1000)
abline(v=20,col="blue", lwd=3)
abline(v=45,col="blue", lwd=3)

hist(unlist(CG_locations), breaks=100)
abline(v=500, col="blue", lwd=3)
```


# For best matches, find probability of given gene regulated

```{r}
best_denovo <- readRDS("../data/best_denovo.rds")
motif_matches <- readRDS("../data/motif_matches.rds")

# Load Regprobs
HM11_regprobs <- readRDS("../data/motifs/HocomocoV11_regprobs_matrix_core_and_extra.rds")
str(HM11_regprobs)

# load motif categories
motif_groups <- read.table("../data/motifs/motif_results_groups.csv", fill=T, sep = ",", stringsAsFactors=F)
motif_groups$group <- c(NA,1:(nrow(motif_groups)-1))
motif_groups <- data.table(melt(motif_groups, id.vars="group"))[value!=""]


regprobs_matrix <- HM11_regprobs[,best_denovo$Target.ID]
colnames(regprobs_matrix) <- sapply(strsplit(colnames(regprobs_matrix), "_"),function(x) x[1])

regprobs_matrix_all <- HM11_regprobs[,motif_matches[Target.Name %in% motif_groups[!is.na(group)]$value]$Target.ID]
colnames(regprobs_matrix_all) <- sapply(strsplit(colnames(regprobs_matrix_all), "_"),function(x) x[1])

str(regprobs_matrix)
str(regprobs_matrix_all)
```


add custom (new) motifs to matrix 

```{r}

results_custom <- readRDS("../data/motifs/MF_results_custom.rds")
regprobs_matrix_custom <- sapply(results_custom, function(x) x$dt$regprob)
rownames(regprobs_matrix_custom) <- results_custom[[1]]$dt$sequence
#colnames(regprobs_matrix_custom) <- paste0(colnames(regprobs_matrix_custom),"*")

regprobs_matrix_custom <- regprobs_matrix_custom[which(rownames(regprobs_matrix_custom) %in% names(tss_seqs_s)), ]

# add CpG Motif to matrix
regprobs_matrix_custom <- cbind("CpG"=CG_count[rownames(regprobs_matrix_custom)]/max(CG_count),
                                regprobs_matrix_custom)

#regprobs_matrix <- cbind(regprobs_matrix, regprobs_matrix_custom[,c("ATF1*","TAF1*")])
#regprobs_matrix_all <- cbind(regprobs_matrix_all, regprobs_matrix_custom[,c("ATF1*","TAF1*")])

```


```{r}

results$loadings_split <- split_loadings()[rownames(HM11_regprobs),]

loadings_subset <- t(results$loadings[[1]][,rownames(regprobs_matrix)])


```



```{r}
qplot(1:50,tapply(CG_count[names(sort(loadings_subset[,"V5"],T))],
                  cut(seq_along(CG_count[names(loadings_subset[,"V5"])]), 50),
                  function(x) mean(x,na.rm=T)),
      ylab="Mean CpG count",
      xlab="Binned Component 5 loadings")

summary(lm(loadings_subset[,"V5"] ~ CG_count[rownames(loadings_subset)]*(loadings_subset[,"V5"]>0)))
```



# Ranked Component Loading vs Motif Probability

```{r, fig.width=6}
# plot raw associations
# regrank_matrix <- apply(regprobs_matrix_custom, 2, rank)

plot(rank(results$loadings[[1]][5,rownames(regprobs_matrix)]),rank(regprobs_matrix[,"NRF1"]))
plot(rank(results$loadings[[1]][5,rownames(regprobs_matrix)]),rank(regprobs_matrix[,"SP2"]))
plot(rank(results$loadings[[1]][42,rownames(regprobs_matrix)]),rank(regprobs_matrix[,"MYBA"]))
plot(rank(results$loadings[[1]][11,rownames(regprobs_matrix)]),rank(regprobs_matrix[,"SPI1"]))

plot(rank(results$loadings[[1]][17,rownames(regprobs_matrix)]),rank(regprobs_matrix_custom[,"ATF1"]))

```

<!-- In early component (42), genes in component (black) have high-skewed regprobs (conversley genes with high THAP1 regprobs have higher prob of being in component). -->

<!-- In late component (17), genes in component (black), have lower-skewed regprobs (i.e. genes with high THAP1 more likely to not be in this component). -->

<!-- ```{r, fig.width=6} -->

<!-- in42 <- results$loadings[[1]][42,][results$pips[[1]][42,]>0.5] -->
<!-- in42 <- in42[names(in42) %in% rownames(regprobs_matrix)] -->
<!-- in42 <- in42[in42<0] -->

<!-- not42 <- results$loadings[[1]][42,][results$pips[[1]][42,]<0.5] -->
<!-- not42 <- not42[names(not42) %in% rownames(regprobs_matrix)] -->
<!-- not42 <- not42[not42<0] -->

<!-- in17 <- results$loadings[[1]][17,][results$loadings[[1]][17,]>0.7] -->
<!-- in17 <- in17[names(in17) %in% rownames(regprobs_matrix)] -->
<!-- in17 <- in17[in17>0] -->

<!-- not17 <- results$loadings[[1]][17,][results$pips[[1]][17,]<0.5] -->
<!-- not17 <- not17[names(not17) %in% rownames(regprobs_matrix)] -->
<!-- not17 <- not17[not17>0] -->

<!-- plot(density(regprobs_matrix[,"THAP1"][names(results$loadings[[1]][42,rownames(regprobs_matrix)][names(in42)])])) -->
<!-- lines(density(regprobs_matrix[,"THAP1"][names(results$loadings[[1]][42,rownames(regprobs_matrix)][names(not42)])]), col="red") -->

<!-- plot(density(regprobs_matrix[,"ATF1*"][names(results$loadings[[1]][17,rownames(regprobs_matrix)][names(in17)])])) -->
<!-- lines(density(regprobs_matrix[,"ATF1*"][names(results$loadings[[1]][17,rownames(regprobs_matrix)][names(not17)])]), col="red") -->
<!-- ``` -->


<!-- ```{r, fig.width=6} -->

<!-- library(ggridges) -->

<!-- tmp <- sort(results$loadings[[1]][17,]) -->
<!-- tmp <- tmp[names(tmp) %in% rownames(regprobs_matrix)] -->

<!-- bin_factor <- cut(seq_along(tmp), 50, labels = 1:50) -->

<!-- ggplot(data.table(value = regprobs_matrix[names(tmp),"THAP1"], quantile = bin_factor)[quantile %in% seq(0,50,2)], aes(value, colour=quantile)) + -->
<!--   geom_density() + -->
<!--   scale_colour_viridis(discrete = T) + theme(legend.position = "none" ) -->

<!-- ggplot(data.table(value = regprobs_matrix[names(tmp),"THAP1"], quantile = bin_factor), aes(value, quantile)) + -->
<!--   geom_density_ridges(aes(fill = quantile)) + -->
<!--   scale_fill_viridis(discrete = T) + theme(legend.position = "none" ) -->

<!-- ggplot(data.table(value = regprobs_matrix[names(tmp),"THAP1"], quantile = bin_factor), aes(quantile, value)) + -->
<!--   geom_boxplot() + -->
<!--   scale_fill_viridis(discrete = T) + theme(legend.position = "none" ) -->

<!-- tmp <- sort(results$loadings[[1]][42,]) -->
<!-- tmp <- tmp[names(tmp) %in% rownames(regprobs_matrix)] -->

<!-- ggplot(data.table(value = regprobs_matrix[names(tmp),"THAP1"], quantile = bin_factor)[quantile %in% seq(0,50,2)], aes(value, colour=quantile)) + -->
<!--   geom_density() + -->
<!--   scale_colour_viridis(discrete = T) + theme(legend.position = "none" ) -->
<!-- ``` -->

<!-- ```{r, fig.width=6} -->


<!-- tmp <- sort(results$loadings[[1]][42,]) -->
<!-- tmp <- tmp[names(tmp) %in% rownames(regprobs_matrix)] -->
<!-- tmp <- data.table(value = regprobs_matrix[names(tmp),"THAP1"], quantile = bin_factor) -->
<!-- tmp <- sapply(1:50, function(x) density(tmp[quantile==x]$value)$y) -->

<!-- ggplot(melt(data.table(tmp, id=1:512), id.vars = "id"), aes( -->
<!--   x=as.numeric(id),  -->
<!--   z=as.numeric(value),  -->
<!--   y=as.numeric(variable), -->
<!--   color=variable)) + -->
<!--   theme_void() + -->
<!--   axes_3D(theta=-30, phi=20) + -->
<!--   stat_3D(theta=-30, phi=20, geom="path")+ -->
<!--   scale_colour_viridis(discrete = T)  -->

<!-- tmp <- sort(results$loadings[[1]][17,]) -->
<!-- tmp <- tmp[names(tmp) %in% rownames(regprobs_matrix)] -->
<!-- tmp <- data.table(value = regprobs_matrix[names(tmp),"THAP1"], quantile = bin_factor) -->
<!-- tmp <- sapply(1:50, function(x) density(tmp[quantile==x]$value)$y) -->

<!-- ggplot(melt(data.table(tmp, id=1:512), id.vars = "id"), aes( -->
<!--   x=as.numeric(id),  -->
<!--   z=as.numeric(value),  -->
<!--   y=as.numeric(variable), -->
<!--   color=variable)) + -->
<!--   theme_void() + -->
<!--   axes_3D(theta=-30, phi=20) + -->
<!--   stat_3D(theta=-30, phi=20, geom="path")+ -->
<!--   scale_colour_viridis(discrete = T)  -->



<!-- # does setting PIP<0.5 to 0 make a difference? - Not really -->
<!-- # tmp <- results$loadings[[1]][46,] -->
<!-- # tmp2 <- tmp -->
<!-- # tmp2[results$pips[[1]][46,]<0.5] <- 0 -->
<!-- # SDAtools::genome_loadings(tmp2) -->

<!-- ``` -->


# Binned Component Loading vs Mean Motif Probability

```{r, fig.width=6}
# For a give component, display binned associations

library(testisAtlas)

binned_regprobs_custom <- binned_mean_regprob(regprobs = regprobs_matrix_custom)

plot_component_regprobs(11, binned_regprobs_custom)
plot_component_regprobs(5, binned_regprobs_custom)
plot_component_regprobs(42, binned_regprobs_custom)
plot_component_regprobs(28, binned_regprobs_custom)
plot_component_regprobs(30, binned_regprobs_custom)
plot_component_regprobs(17, binned_regprobs_custom)
```

compare to the mean loadings per bin of loading

```{r}
bin_factor <- cut(seq_along(regprobs_matrix[,1]), 50)

tmp <- sort(results$loadings[[1]][30,])
tmp <- tmp[names(tmp) %in% rownames(regprobs_matrix)]
plot(tapply(tmp, bin_factor, mean), main="30")

tmp <- sort(results$loadings[[1]][5,])
tmp <- tmp[names(tmp) %in% rownames(regprobs_matrix)]
plot(tapply(tmp, bin_factor, mean), main="5")

tmp <- sort(results$loadings[[1]][17,])
tmp <- tmp[names(tmp) %in% rownames(regprobs_matrix)]
plot(tapply(tmp, bin_factor, mean), main="17")

tmp <- sort(results$loadings[[1]][42,])
tmp <- tmp[names(tmp) %in% rownames(regprobs_matrix)]
plot(tapply(tmp, bin_factor, mean), main="42")
```

# Mean regprob by component

```{r, fig.width=10}



library(ComplexHeatmap)
library(RColorBrewer)

plot_meanprob_bycomp(binned_regprobs_custom)

Heatmap(t(binned_regprobs_custom["SP2",,component_order]),
        col=rev(colorRampPalette(brewer.pal(11, "RdYlBu"))(100)),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "SP2")

Heatmap(t(binned_regprobs_custom["RFX2",,component_order]),
        col=rev(colorRampPalette(brewer.pal(11, "RdYlBu"))(100)),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "RFX2")

Heatmap(t(binned_regprobs_custom["GABPA",,component_order]),
        col=rev(colorRampPalette(brewer.pal(11, "RdYlBu"))(100)),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "GABPA")

Heatmap(t(binned_regprobs_custom["MYBA",,component_order]),
        col=rev(colorRampPalette(brewer.pal(11, "RdYlBu"))(100)),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "MYBA")

Heatmap(t(binned_regprobs_custom["ATF1",,component_order]),
        col=rev(colorRampPalette(brewer.pal(11, "RdYlBu"))(100)),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_title = "ATF1*")

mean_prob_matrix <- cbind(binned_regprobs_custom[,50,],binned_regprobs_custom[,1,])
colnames(mean_prob_matrix) <- colnames(results$loadings_split)

mean_prob_matrix <- t(mean_prob_matrix[,ordering[-half_exclusions]])
rownames(mean_prob_matrix) <- paste(rownames(mean_prob_matrix), component_names[-half_exclusions])

NMF::aheatmap(scale(mean_prob_matrix), Rowv = NA, hclustfun = "ward.D2")

```


# Correlation of Motif Probability and Gene Loadings

For all motifs

```{r, fig.width=11}
library(RColorBrewer)

vvv1 <- correlate_regprobs(regprobs = regprobs_matrix_all)

vvv1 <- vvv1[ordering[-half_exclusions],]
rownames(vvv1) <- paste(rownames(vvv1), component_names[-half_exclusions]) # paste(rownames(vvv2), c(rbind(names(component_labels), names(component_labels)))[-half_exclusions])

pdf(height=15, width=25, file="../results/motifs_fig_HCMC2.pdf")
aheatmap(asinh(vvv1), scale="none", Rowv = NA, cexCol=1.3,
         color=rev(colorRampPalette(brewer.pal(11, "RdYlBu"))(100)))
dev.off()
```

For selected motifs

```{r}
vvv3 <- correlate_regprobs(regprobs = regprobs_matrix_custom)

vvv3 <- vvv3[ordering[-half_exclusions],]
rownames(vvv3) <- paste(rownames(vvv3), component_names[-half_exclusions])

motif_order <- c("SP2","ETV5","CpG","TAF1","ZN143","NRF1","GABPA","NFYA","ZFP42","MYBA","MLX","CREM","RFX2","TBP","STAT1","SPI1","ATF1")

vvv3 <- vvv3[,motif_order]

motif_gene_names <- c("Sp2","Etv5","CpG","Taf1","Zfp143","Nrf1","Gabpa","Nfya","Zfp42","Mybl1","Mlx","Crem","Rfx2","Tbp","Stat1","Spi1","Crem-t")
colnames(vvv3) <- motif_gene_names

library(ComplexHeatmap)

results_custom_motifs <- sapply(results_custom, function(x) x$scoremat)
hascpg <- c("CpG"=NA,sapply(results_custom_motifs, function(x) MotifFinder::submotif(exp(x))))

meth_sens = c("SP2"="No","ETV5"="Yes","CpG"="NA","TAF1"="No","ZN143"="No","NRF1"="Yes","GABPA"="Yes","NFYA"="Yes",
                      "ZFP42"="Yes","MYBA"="No","MLX"="Yes","CREM"="Yes","RFX2"="No","TBP"="No","STAT1"="No","SPI1"="Yes","ATF1"="NA")

hascpg["CpG"] <- max(hascpg, na.rm = T)

dt1 <- data.table("P(CpG)>0.25:" = hascpg[motif_order]>0.25,
                  "P(CpG):" = hascpg[motif_order],
                  "Meth Sensitive?:" = meth_sens[motif_order])

ha1 = HeatmapAnnotation(df = dt1,
    col = list("Has CpG?:"=c("Yes"="purple","No"="grey"),
               "Meth Sensitive?:"=c("Yes"="dark green","No"="grey","NA"="grey28"),
               "P(CpG)>0.25:"=c("TRUE"="purple","FALSE"="grey"),
               "P(CpG):"=circlize::colorRamp2(c(min(hascpg), max(hascpg)), c("white", "red"))), #colorRamp2(seq(0, 1, by = 0.25), viridisLite::viridis(5, direction = -1))
               annotation_legend_param = list(grid_height = unit(3, "mm"),
                                              labels_gp = gpar(fontsize = 7),
                                              legend_direction="horizontal",
                                              ncol = 3,
                                              title_position = "lefttop")
)

pdf(height=5, width=10, file="../results/motifs/motifs_fig_DN_asinh.pdf")
draw(
Heatmap(asinh(vvv3),
        col = rev(colorRampPalette(bias=1.1, brewer.pal(11, "RdYlBu"))(100)), #viridisLite::cividis(200),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_order = motif_gene_names,
        row_order = 1:nrow(vvv3),
        top_annotation = ha1,
        heatmap_legend_param = list(legend_direction = "horizontal", title = "asinh(t-value)"),
        row_names_gp = gpar(fontsize = 3),
        show_heatmap_legend = T,
        top_annotation_height = unit(5, "mm")),
split = rep(c("1) Somatic", "2) Diploid", "3) Haploid"), c(23,33,19)),
row_dend_gp = gpar(fontsize = 3),
annotation_legend_side = "bottom",
heatmap_legend_side = "bottom")
dev.off()


# version without asinh thresholding
pdf(height=5, width=10, file="../results/motifs/motifs_fig_DN_raw.pdf")
draw(
Heatmap(vvv3,
        col = rev(colorRampPalette(bias=1.1, brewer.pal(11, "RdYlBu"))(100)), #viridisLite::cividis(200),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_order = motif_gene_names,
        row_order = 1:nrow(vvv3),
        top_annotation = ha1,
        heatmap_legend_param = list(legend_direction = "horizontal", title = "asinh(t-value)"),
        row_names_gp = gpar(fontsize = 3),
        show_heatmap_legend = T,
        top_annotation_height = unit(5, "mm")),
split = rep(c("1) Somatic", "2) Diploid", "3) Haploid"), c(23,33,19)),
row_dend_gp = gpar(fontsize = 3),
annotation_legend_side = "bottom",
heatmap_legend_side = "bottom")
dev.off()

```

For selected motifs - regprobs using Hocomoco not denovo PWM

```{r}
results$loadings_split <- split_loadings()

vvv2 <- correlate_regprobs(regprobs = regprobs_matrix[,motif_order[-3]])

vvv2 <- vvv2[ordering[-half_exclusions],]
rownames(vvv2) <- paste(rownames(vvv2), component_names[-half_exclusions])

library(ComplexHeatmap)

pdf(height=5, width=10, file="../results/motifs_fig_HCMC.pdf")
draw(
Heatmap(asinh(vvv2),
        col=rev(colorRampPalette(brewer.pal(11, "RdYlBu"))(100)),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        row_order = 1:nrow(vvv2),
        row_names_gp = gpar(fontsize = 3),
        show_heatmap_legend = F,
        top_annotation_height = unit(2, "mm")),
split = rep(c("1) Somatic", "2) Diploid", "3) Haploid"), c(23,33,19)),
row_dend_gp = gpar(fontsize = 3),
annotation_legend_side = "bottom")
dev.off()
```

# More motifs

```{r}

# select motifs, HCMC regprobs
save_regprobs_heatmap(regprobs_matrix, "../results/motifs/motifs_fig_DN_HCMCregprobs_subset.pdf", width=5)

# all denovo motifs, but using hocomoco regprobs
save_regprobs_heatmap(regprobs_matrix_all, "../results/motifs/motifs_fig_DN_HCMCregprobs.pdf", width=10)

# All denovo motifs, denovo regprobbs
save_regprobs_heatmap(regprobs_matrix_custom, "../results/motifs/motifs_fig_DN_All.pdf", width=10)

```

# All Hocomoco Motifs

```{r}
regprobs_matrix_core <- sapply(results_core, function(x) x$dt$regprob)
rownames(regprobs_matrix_core) <- results_core[[1]]$dt$sequence

save_regprobs_heatmap(regprobs_matrix_core, "../results/motifs/motifs_fig_HM_core.pdf", width=100)

save_regprobs_heatmap(HM11_regprobs[,1:300], "../results/motifs/motifs_fig_HM11_regprobsA.pdf", width=100)
save_regprobs_heatmap(HM11_regprobs[,301:600], "../results/motifs/motifs_fig_HM11_regprobsB.pdf", width=100)
save_regprobs_heatmap(HM11_regprobs[,601:900], "../results/motifs/motifs_fig_HM11_regprobsC.pdf", width=100)
save_regprobs_heatmap(HM11_regprobs[,901:1276], "../results/motifs/motifs_fig_HM11_regprobsD.pdf", width=100)
```


# All Denovo motifs from single components


```{r}

results_C2N <- readRDS("../data/motifs/MF_results_C2N_V3.rds")
regprobs_matrix_C2N <- sapply(results_C2N, function(x) x$dt$regprob)
rownames(regprobs_matrix_C2N) <- results_C2N[[1]]$dt$sequence

regprobs_matrix_C2N <- regprobs_matrix_C2N[which(rownames(regprobs_matrix_C2N) %in% names(tss_seqs_s)), ]

save_regprobs_heatmap(regprobs_matrix_C2N, "../results/motifs/motifs_fig_DN_C2N.pdf", width = 7)

#####


results_C5N <- readRDS("../data/motifs/MF_results_C5N_V3.rds")
regprobs_matrix_C5N <- sapply(results_C5N, function(x) x$dt$regprob)
rownames(regprobs_matrix_C5N) <- results_C5N[[1]]$dt$sequence

regprobs_matrix_C5N <- regprobs_matrix_C5N[which(rownames(regprobs_matrix_C5N) %in% names(tss_seqs_s)), ]

save_regprobs_heatmap(regprobs_matrix_C5N, "../results/motifs/motifs_fig_DN_C5N.pdf", width = 7)

######

results_C11P <- readRDS("../data/motifs/MF_results_C11P_V3.rds")
regprobs_matrix_C11P <- sapply(results_C11P, function(x) x$dt$regprob)
rownames(regprobs_matrix_C11P) <- results_C11P[[1]]$dt$sequence

regprobs_matrix_C11P <- regprobs_matrix_C11P[which(rownames(regprobs_matrix_C11P) %in% names(tss_seqs_s)), ]

save_regprobs_heatmap(regprobs_matrix_C11P, "../results/motifs/motifs_fig_DN_C11P.pdf", width = 7)

######

results_C3P <- readRDS("../data/motifs/MF_results_C3P_V3.rds")
regprobs_matrix_C3P <- sapply(results_C3P, function(x) x$dt$regprob)
rownames(regprobs_matrix_C3P) <- results_C3P[[1]]$dt$sequence

regprobs_matrix_C3P <- regprobs_matrix_C3P[which(rownames(regprobs_matrix_C3P) %in% names(tss_seqs_s)), ]

save_regprobs_heatmap(regprobs_matrix_C3P, "../results/motifs/motifs_fig_DN_C3P.pdf", width = 7)


######

results_novel_A <- readRDS("../data/motifs/MF_results_novel_A.rds")
regprobs_matrix_novel_A <- sapply(results_novel_A, function(x) x$dt$regprob)
rownames(regprobs_matrix_novel_A) <- results_novel_A[[1]]$dt$sequence

save_regprobs_heatmap(regprobs_matrix_novel_A, "../results/motifs/motifs_fig_DN_novelA.pdf", width=15)



results_novel_B <- readRDS("../data/motifs/MF_results_novel_B.rds")
regprobs_matrix_novel_B <- sapply(results_novel_B, function(x) x$dt$regprob)
rownames(regprobs_matrix_novel_B) <- results_novel_B[[1]]$dt$sequence

save_regprobs_heatmap(regprobs_matrix_novel_B, "../results/motifs/motifs_fig_DN_novelB.pdf", width=15)


```

# Distribution of Positions relative to TSS

```{r}
whichpos_matrix_custom <- sapply(results_custom, function(x) x$dt[order(-regprob)][1:3000]$whichpos)

whichpos_dt <- melt(data.table(t(whichpos_matrix_custom[,motif_order[c(-1)]]), keep.rownames = T), id.vars = "rn")

whichpos_dt[,rn := factor(rn, levels=motif_order[c(-1)])]

ggplot(whichpos_dt, aes(value-500)) +
  geom_density() +
  facet_wrap(~rn, scales="free_y") +
  geom_vline(xintercept = 0, col='red') + 
  xlab("Distance from TSS") + 
  ylab("Relative Density")
```

# Attempt: expression due to single TF

```{r}


common_genes <- rownames(regprobs_matrix_custom)[rownames(regprobs_matrix_custom) %in% colnames(data)]

tmp <- data[,common_genes] * matrix(rep(t(regprobs_matrix_custom[common_genes,"CREMT",drop=F]),nrow(data)),nrow=nrow(data),byrow = T)
tmp2 <- Matrix::rowMeans(tmp)
ggplot(data.table(Stra8=tmp2, cell=names(tmp2))[datat,on="cell"], aes(Tsne1_QC1, Tsne2_QC1, colour=Stra8)) + geom_point(stroke=0, size=1) + scale_colour_viridis(direction = -1)

```


